STAT 133 Final Project: Visualizing Berkeley Police Data

### 2016-08-09 final_paper/finalpaper.Rmd
updates stop data

# histogram of stop data by race and hour:
histogram_stoptime1 <- ggplot(stops, 
                              aes(x = Hour, 
                                  fill = Race)) + 
  geom_bar(aes(y = ..count..)) +
  facet_grid(~Race)+
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  labs(title = "Number of people each race in every hour",
       y = "Count") +
  guides(fill = FALSE)+
  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 14),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

@@

stop_percentage2 <- ggplot(age_hour , aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  facet_wrap(~AgeRange)+
labs(title = "Probability of BPD stop for a \nspecific age range in a specific hour",
       x= "Hour",
       y = "Density") +
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) +
  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 10),
        axis.text.x=element_text(colour="black", size = 7), 
        axis.text.y=element_text(colour="black", size = 7)) +
  guides(fill = FALSE)

@@

stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  facet_wrap(~Reason)+
  labs(title = "Probability of BPD stop for a \nspecific age range in a specific reason",
       x= "Hour",
       y = "Density") +
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  geom_text(aes(x=AgeRange,
               y=percentage,
               label = paste0(percentage*100, "%")), 
            size=3,
            colour= "#000000")+
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) +
  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 10),
        axis.text.x=element_text(colour="black", size = 7), 
        axis.text.y=element_text(colour="black", size = 7)) +
  guides(fill = FALSE)

@@

## PLEASE CHANGE THE TABLE 
#### LOVE JENNY

A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, fig.align="center"}
grid.arrange(bargraph1, bargraph2, ncol = 2)
```

@@

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
map_reasons1
```

!!!!!!!!!!!!!!!!!!!!!INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE (Use percentage, not count) !!!!!!!!!!!!!!!!!!!!!!!!

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest
```

###  Findings 
1. The probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%.

2. The probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%.

3. An interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race. Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_percentage3
```
###  Findings 
1. Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%.

2. White people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%.

3. Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_percentage2 
```
### Findings 
1. People aged from 18 to 29 are obviously liable to commit incidents at night. The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%.

2. People aged greater than 40 are obviously liable to commit incidents in the daytime. The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%.

3. Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest_age
```
### Jenny Findings 
1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.

2. Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range.

3. The conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range. Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.

```{r, fig.height = 10, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest_gender 
```
### Jenny Findings 
1. Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest.

2. Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 5, fig.width = 10, fig.align="center"}
#### INCLUDE BAR GRAPH & TABLE OF POLICE STOPS BY REASON HERE(Use percentage, not count)!!!!!!!!!!!!!!!!!!!!!!!!
grid.arrange(bararrest, barjail, ncol = 2)
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 12, fig.height=3, fig.align="center"}
histogram_jail1
histogram_arrest1
# I decided to take these out of the multi.plot function because I was trying to adjust the format a certain way, I'm sorry!
```

@@

>>>>

# histogram of stop data by race and hour:

histogram_stoptime1 <- ggplot(stops, 
                              aes(x = Hour, 
                                  fill = Race)) + 
  geom_bar(aes(y = ..count..)) +
  facet_grid(~Race)+
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  labs(title = "Number of people each race in every hour",
       y = "Count") +
  guides(fill = FALSE)+
  theme(plot.title = element_text(size = 14, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 14),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

@@

stop_percentage2 <- ggplot(age_hour , aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  facet_wrap(~AgeRange, nrow = 1)+
labs(title = "Probability of BPD stop for a \nspecific age range in a specific hour",
       x= "Hour",
       y = "Density") +
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) +
  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 10),
        axis.text.x=element_text(colour="black", size = 7), 
        axis.text.y=element_text(colour="black", size = 7)) +
  guides(fill = FALSE)

@@

stop_arrest_age <- ggplot(stop_by_percentage_age, aes(x = AgeRange, fill = AgeRange)) + 
  geom_bar(aes(y = percentage), stat="identity") +
  facet_wrap(~Reason, nrow=1)+
  labs(title = "Probability of BPD stop for a \nspecific age range in a specific reason",
       x= "Hour",
       y = "Density") +
  scale_fill_manual(values=c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")) +
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  geom_text(aes(x=AgeRange,
               y=percentage,
               label = paste0(percentage*100, "%")), 
            size=3,
            colour= "#000000")+
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) +
  theme(plot.title = element_text(size = 12, family = "Georgia", face = "bold"), 
        text=element_text(family="Times New Roman", size = 10),
        axis.text.x=element_text(colour="black", size = 7), 
        axis.text.y=element_text(colour="black", size = 7)) +
  guides(fill = FALSE)

@@

## PLEASE mention that we delete tHE TABLE 
#### LOVE JENNY

A visualization of the groups of people stopped in proportion to their population is provided by the bar chart below.
```{r, message = F, echo = FALSE, fig.width = 10, fig.height = 3, warning = FALSE, fig.align="center"}
grid.arrange(bargraph1, bargraph2, ncol = 2)
```

@@

```{r, fig.height = 5, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
map_reasons1
```

## PLEASE EDIT THE ANALYSIS 
###  LOVE JENNY

We also analyzed the preference of Berkeley PD when arresting people. We calculated the conditional probability of arrested given a specific reasion and given a specific race/ age range/ gender. That is to say, we calculated the probability of a person will be arrested by Berkeley PD if he was stopped by a specific reason and was of a specific race/ age range/ gender.

```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest
```
## PLEASE EDIT THE ANALYSIS 
###  LOVE JENNY 
From the above plots, we can tell that the probability of arrested in a stop with the reason Traffic is much lower than other reasons. The average conditional probability of arrested given reason is traffic is 1.58%. Additionally, the probability of arrested in a stop with the reason Wanted is much higher than other reason. The average conditional probability of arrested given reason is wanted is 33.33%. Lastly, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and race Asian is 75%, which is much higher than that of any other race. Asian people is much more liable to be arrested by Berkeley Police Department during Probation or Parole if stopped by the BPD.

```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest_age
```

## FINDINGS
###  LOVE JENNY 
Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest. Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole and age range 0-18 is 67.56%, which is much higher than that of any other age range. Lastly, the conditional probability of arrested given person involved aged from 0 to 18 in a stop given any reason is more than that of given person with any other age range. Therefore, in a stop, teenagers is more liable to be arrested by Berkeley Police Department.

```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_arrest_gender
```

## FINDINGS
####  LOVE JENNY 
Similar to the conclusions above, the probability of arrested in a stop with the reason Traffic is the lowest and wanted is the highest. Again similar to the conclusion above, an interesting fact is that the conditional probability of arrested in a stop with the reason Probation or Parole of female is 50.00%, which is much higher than that of male, which is 20.11%.

```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_percentage3
```
##  Findings 
#### LOVE JENNY
Black people are obviously liable to commit incidents at night. The average ratio of black people to all people stoped by the Berkeley Police Department at night is about 40%. On the other hand, white people are obviously liable to commit incidents in the daytime. The average ratio of white people to all people stoped by the Berkeley Police Deparment during the day is about 45%. While the Ratio of incidents commited by Asian people, Hispanic people and other people fluctuate during the daytime and the night, with an average ratio of 8%, 11% and 9% respectively.

```{r, fig.height = 3, fig.width = 10, echo=FALSE, warning = FALSE, message = FALSE, fig.align = "center"}
stop_percentage2
```

## Findings 
#### LOVE JENNY  
People aged from 18 to 29 are obviously liable to commit incidents at night. The average ratio of people aged from 18 to 29 to all people stoped by the Berkeley Police Department at night is greater than 40%. People aged greater than 40 are obviously liable to commit incidents in the daytime. The average ratio of people aged greater than 40 to all people stoped by the Berkeley Police Deparment during the day is greater than 40%. While Ratio of incidents commited by people aged between 0 and 18 and people aged from 30 to 39 fluctuates during the daytime and the night, with an average ratio of 2.5% and 25% respectively.

@@

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 3, fig.width = 10, fig.align="center"}
grid.arrange(bararrest, barjail, ncol = 2)
```

@@

## PLEASE MENTION THE HOMELESS PEOPLE. 
#### LOVE JENNY
